Smart Drill-Down: Interactive Summarization of Relational Tables

### 2014-11-18 TableSummarization/Paper/algorithms.tex
Addressed Comments. Added NP-Hardness proof

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r,R)$$
\end{problem}

\agp{not essential to mention here.}
\sout{$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.}

\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\squishlist
\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\squishend

\agp{not essential to mention here}\sout{We will describe a greedy algorithm for finding the approximately highest scoring rule-list.}

First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.
\agp{mention that this is independent of MCount. Also give intuition--this is because
we can transfer tuples from one rule to a better rule adn only increase score.}

@@

\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. This gives us 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a weight function $W$, and a number $k$, to find the set $R$ of $k$ rules such that if $R^{\prime}$ is the rule-list obtained by ordering rules in $R$ in decreasing weight order, then the following quantity is maximised:
$$\text{Score}(R) = \text{Score}(R^{\prime}) = \sum_{r \in R^{\prime}}W(r)MCount(r,R)$$
\end{problem}

\agp{Confusing since you're conflating lists and sets. Explicitly define the score of a set of rules
is the score of the best rule list (and in the equation above, have max over all R' where R' is an ordered
version of S)}

\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now show an approximation algorithm for Problem~\ref{prob:optimal-rule-set}.

\subsection{Algorithm Overview}\label{sec:alg-overview}
\agp{Give a high-level perspective of the algorithm; 
Basics: greedy approximation + a-priori; 
Advanced: Sampling for larger datasets: comprising two parts
how to best use samples, and how to select samples}

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
\agp{can we show hardness in data or schema complexity?}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

@@

\subsection{Sampling for big tables}

\agp{Here also, there is a need for some organization. There are clearly
some good ideas here, but you need to give a high-level picture first before
diving in.}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if it does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

@@

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be. We provide a way to find reasonable values of $minSS$ later.

@@

For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples. Call this quantity $S(r_1, r_2)$. We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as $\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$. If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table. We wish to maximise the probability that we can respond to the next user expansion without making another pass.

\agp{not clear what problem you're trying to solve here. consider raising it up as a subsection
and giving it a name. }
Thus formally our problem is :
Choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
where the $I$'s are indicator variables, subject to :
$$\sum_{r \in U} n_r < \text{capacity}$$
This problem is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.

>>>>

\begin{problem}\label{prob:optimal-rule-list}
Given a table $T$, a monotonic weight function $W$, and a number $k$, to find the list $R$ of $k$ rules that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)MCount(r,R)$$
\end{problem}

\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\begin{enumerate}
\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\end{enumerate}

First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight. Note that the weight of a rule is independent of its $MCount$. The $MCount$ of a rule is the number of tuples that have been `assigned' to it, and each tuple assigned to rule $r$ contributes $W(r)$ to the total score. Thus if the rules are not in decreasing order by weight in a rule list $R$, then switching the order of rules in $R$ transfers some tuples from a lower weight rule to a higher weight rule, which can increase total score.

@@

\noindent Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, we can define Score for a set of rules as follows:

\begin{definition}\label{def:set-score}
Let $R$ be a set of rules. Then the Score of $R$ is defined by
$$\text{Score}(R) = \text{Score}(R^{\prime})$$
where $R^{\prime}$ is the list of rules obtained by ordering the rules in $R$ in decreasing order by weight.
\end{definition}

This gives us a reduced version of Problem~\ref{prob:optimal-rule-list}: 
\begin{problem}\label{prob:optimal-rule-set}
Given a table $T$, a monotonic weight function $W$, and a number $k$, find the set $R$ of $k$ rules which maximizes Score($R$) as defined in Definition~\ref{def:set-score}.
\end{problem}

\noindent The reduction from Problem~\ref{prob:optimal-rule-list} to Problem~\ref{prob:optimal-rule-set} is clear. We now first show that Problem~\ref{prob:optimal-rule-set}, and consequently Problem~\ref{prob:optimal-subrule-list} and Problem~\ref{prob:optimal-rule-list} are NP-Hard, and then present an approximation algorithm for solving Problem~\ref{prob:optimal-rule-set}.

\subsection{NP-Hardness for Problem~\ref{prob:optimal-rule-set}}
We reduce the well known NP-Hard problem called the {\em Maximum Coverage Problem} to a special case of Problem~\ref{prob:optimal-rule-set}, which will prove NP-Hardness for Problem~\ref{prob:optimal-rule-set}. The Maximum Coverage Problem is given below: 
\begin{problem}\label{prob:maximum-coverage}
Given a universe set $U$, an integer $k$, and a set $S = \left\lbrace S_1, S_2, ... S_m \right\rbrace$ of subsets of $U$ (so each $S_i \subset U$), find $S^{\prime} \subset S$ such that $|S^{\prime}| = k$, which maximises $\text{Coverage}(S^{\prime}) = |\bigcup_{s \in S^{\prime}} s|$.
\end{problem}
Thus the maximum coverage problem is to find a set of $k$ of the given subsets of $U$ whose union `covers' as much of $U$ as possible. We now reduce an instance of the Maximum Coverage Problem (with parameters $U, k, S$) to an instance of Problem~\ref{prob:optimal-rule-set}.

Then we construct the table $T$ as follows: $T$ has $|U|$ tuples and $|S|$ columns. Column $i$ of $T$ corresponds to set $S_i \in S$ and each tuple of $T$ corresponds to one $u \in U$. For each $u \in U$, the tuple $t_u$ corresponding to $u$ defined as follows: For each $S_i \in S$, $$t_u(i) = 1 \textrm{ if } u \in S_i \textrm{ else } t_u(i) = 0$$

Define the weight function $W$ as follows:
$$W(r) = \textrm{min}(1, |\left\lbrace i \in \left\lbrace 1, 2, ... m \right\rbrace : r(i) = 1 \right\rbrace|)$$
Thus $W$ gives weight $0$ to rules that do not have value $1$ in any column, and weight $1$ to all other rules. An optimal rule set for weight function $W$ on table $T$ need only contain rules having a single $1$ value and no $0$ values, because adding more values can only reduce $MCount$, without increasing weight.

Then a solution to the instance of Problem~\ref{prob:optimal-rule-set} with table $T$, weight function $W$, and required number of rules $k$ gives us a solution to the instance of Problem~\ref{prob:maximum-coverage} as follows: For each rule $r$ in the solution of the Problem~\ref{prob:optimal-rule-set} instance, if it has value $1$ in column $i$ and $\star$ in all other columns, then add $S_i$ to the solution set of the Problem~\ref{prob:maximum-coverage} instance. This completes our reduction.

\subsection{Algorithm Overview}\label{sec:alg-overview}
We now present our algorithms for approximating the solution to Problem~\ref{prob:optimal-rule-set}. The problem consists of finding a set of given size $k$, that maximises Score. We first show that the Score function is {\em submodular}, and hence an approximately optimal set can be obtained using a greedy algorithm. We start with an empty rule set $R$, and then run the greedy algorithm for $k$ steps, where in each step, we add the rule $r$ to $R$ that maximises Score($R \cup \left\lbrace r \right\rbrace$). In order to find the rule $r$ to add in each step, we need to measure $Score(R \cup \left\lbrace r \right\rbrace) - Score(R)$ for each $r$. This is done in several passes over the table, using ideas from the a priori algorithm~\cite{apriori} for frequent itemset mining.

In addition, the greedy algorithm can be run over a sample of the table instead of over the entire table, to reduce computation time for large tables. We later describe a scheme for maintaining multiple samples in memory and using them to improve response time for different drill down operations performed by the user.

\subsection{Greedy Approximation Algorithm}\label{sec:greedy-approx}
We will now show that the Score function over sets of rules has a property called {\em submodularity}, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

@@

\subsection{Sampling for big tables}
\subsubsection{Sampling Overview}
We now describe our sampling schemes for improving the running time of our algorithm on tables that are too large to fit in main memory. First in Section~\ref{sec:sample-using}, we motivate the need for sampling. Then we describe a component of our system, called the {\em SampleHandler}, which is responsible for creating and maintaining samples of the table in memory, subject to user specified memory constraints. The SampleHandler maintains multiple samples corresponding to different parts of the table, which can be used depending on which rule the user decides to expand next. Then in Section~\ref{sec:sampling_algorithms}, we describe ways to efficiently allocate memory to different samples, so as to maximize the probability that we can respond to the next user operation without accessing the hard disk. Finally, we mention some additional optimizations we can make, and heuristics for setting the minimum sample size required from the SampleHandler.

\subsubsection{Using samples to respond to user operations}\label{sec:sample-using}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if it does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

@@

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (both specified by the user).
The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be. We later provide a way to find reasonable values of $minSS$.

@@

For each pair of rules $r_1, r_2 \in U$ such that $r_1$ is a sub-rule of $r_2$, we can estimate the ratio of selectivities of $r_1$ and $r_2$ using existing samples. Call this quantity $S(r_1, r_2)$. We define $S(r_1, r_2)$ to be $0$ if $r_1$ is not a sub-rule of $r_2$. Then if we have an $n_r$ sized sample with filter $r$ for each $r \in U$, the expected number of tuples covered by $r^{\prime} \in L$ is given by $$\sum_{r \in U} S(r, r^{\prime})n_r$$ For any $r^{prime}$, define the function {\em ess} (for `effective sample size') as below: 
\begin{definition}\label{def:ess}
$$\text{ess}(r^{\prime}) = \sum_{r \in U} S(r, r^{\prime})n_r$$ 
\end{definition}
If $\text{ess}(r^{\prime}) \geq \text{minSS}$ then if the user expands $r^{\prime}$, we display the next rule list using \textbf{Find} or \textbf{Combine}, instead of having to make another pass through the table. We wish to maximise the probability that we can respond to the next user expansion without making another pass. We now formally define our problem below:
\begin{problem}\label{prob:sample-sizes}
Given a tree of rules $U$ with leaves $L$, a probability distribution $p$ over $L$, an integer `capacity', and selectivity ratio $S(r_1, r_2)$ for each $r_1, r_2 \in U$, 
choose an integer $n_r \geq 0$ for each $r \in U$ so as to $\textrm{maximize}$ :
$$\sum_{r^{\prime} \in L} p_{r^{\prime}}I_{\text{ess}(r^{\prime}) \geq \text{minSS}}$$
where the $I$'s are indicator variables, subject to :
$$\sum_{r \in U} n_r < \text{capacity}$$
\end{problem}
Problem~\ref{prob:sample-sizes} is non-linear and non-convex because of the indicator variables. If the tree $U$ is not too big, then it might be feasible to use an exponential algorithm to solve the problem.

### 2014-11-18 TableSummarization/Paper/exp.tex
Addressed Comments. Added NP-Hardness proof

\section{Experiments}\label{sec:experiments}
\subsection{Dataset}
The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

The columns (in order) are as follows:
(Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, How long have you lived in the San Fran./Oakland/San Jose Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home).

@@

\subsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

\subsection{Weighting functions}
Our system can display optimal rule summaries using any monotone weighting function. By default, we assign a rule weight equal to it's size. In this section, we try a couple of other weighting functions.

@@

\subsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$. In this section, we study the effect of varying $m_w$ on the speed of our algorithm. Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

@@

\subsection{Effects of $minSS$}
We now study the effects of the sampling parameter $minSS$. This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.   Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.

@@

The computation time increases approximately linearly with $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, which takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$.

>>>>

\section{Experiments}\label{sec:experiments}

\begin{comment} ask about below:
Additional experiments:
Maybe also good to fix mw, and show the performance with or without sampling,
and with or without a-priori (one rule per phase, for instance).
Also, would be good to see performance on increasing dataset size.
\end{comment}

\subsection{Dataset}
The dataset we use is a marketing dataset (MD), which contains demographic information about potential customers. The source of the dataset is Impact Resources, Inc., Columbus, OH (1987). A total of $N=9409$ questionnaires containing $502$ questions were 
filled out by shopping mall customers in the San Francisco Bay area. This dataset is an extract from this survey. Each tuple in the data table describes a single person. There are $14$ columns, each of which is a demographic attribute, such as annual income, Gender, marital status, age, education, and so on. Continuous values, such as income, have been bucketized in the dataset, and thus each column has upto $10$ distinct values.

The columns (in order) are as follows:
Annual Income of Household, Sex, Marital Status, Age, Education, Occupation, Time lived in Bay Area, Dual Incomes?, Persons in Household, Persons in Household under $18$, Householder Status, Type of Home, Ethnic Classification, Language most spoken in home.

@@

\subsection{Qualitative Study}
We first perform a qualitative study of our Smart Drill Down system. We observe the effects of various user interface operations on our dataset, and then try out a few different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

@@

\subsubsection{Weighting functions}
Our system can display optimal rule summaries using any monotonic weighting function. By default, we assign a rule weight equal to it's size. In this section, we try out a couple of other weighting functions.

@@

\subsection{Quantitative Study}
The performance of our algorithm depends on various parameters, such as $m_w$ (the max weight) and $minSS$ (minimum required sample size). In this section, we study the effects of these parameters on the computation time and accuracy of our algorithm.

\subsubsection{Effects of $m_w$}
Our algorithm for finding the best marginal rule (Algorithm~\ref{algo:best-marginal-rule}) takes an input parameter called $m_w$. The algorithm is guaranteed to find the best marginal rule as long as it's weight is $\leq m_w$, but runs faster for smaller values of $m_w$. In this section, we study the effect of varying $m_w$ on the speed of our algorithm. Our system is run on a Dell XPS L702X laptop with 6GB RAM and an Intel i5 2.30GHz processor.

@@

\subsubsection{Effects of $minSS$}
We now study the effects of the sampling parameter $minSS$. This parameter tells the SampleHandler the minimum sample size on which we run Algorithm~\ref{algo:best-rule-set}.   Higher values of $minSS$ cause our system to use bigger samples, which increases the accuracy of count estimates for displayed rules, but also correspondingly increases computation time.

@@

The computation time increases approximately linearly with parameter $minSS$. This is expected, because Algorithm~\ref{algo:best-rule-set} makes a constant number of passes over the entire sample, and each pass takes time linear in the size of the sample. The percent error decreases approximately as $\frac{1}{\sqrt{minSS}}$, which is again expected because the standard deviation of estimated $Count$ is approximately inversely proportional to the square root of sample size.

### 2014-11-18 TableSummarization/Paper/extensions.tex
Many small changes
small suggestions
Addressed Comments. Added NP-Hardness proof

\section{Extensions}\label{sec:extensions}
\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
Our algorithm assumes that all attributes are categorical in nature. Attributes that have a large domain d to have a smaller tuple count per value, and hence don't appear in rule summaries. Thus our algorithm does not summarise information about numerical attributes.

@@

The distribution of values for the $A$ may not be similar in different parts of the table. For instance, we may create buckets have approximately equal numbers of tuples in the table. But then suppose a user tries to expand a rule $r$. The part of the table covered by $r$ might have a very different distribution of age values. For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment. To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.

\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.

>>>>

\section{Extensions}\label{sec:extensions}
\subsection{Dealing with Numerical Attributes}\label{sec:extensions-numerical}
Our algorithm assumes that all attributes are categorical in nature. Attributes that have a large domain tend to have a smaller tuple count per value, and hence don't appear in rule summaries. Thus our algorithm does not summarise information about numerical attributes.

@@

\begin{comment}
The distribution of values for the $A$ may not be similar in different parts of the table. For instance, we may create buckets have approximately equal numbers of tuples in the table. But then suppose a user tries to expand a rule $r$. The part of the table covered by $r$ might have a very different distribution of age values. For instance, in our MD dataset, if the user looks at people who own a house, their age distribution would be different from that of people who are still renting an apartment. To prevent the buckets from getting a skewed distribution when expanding $r$, we could recompute the bucket ranges every time the user expands a rule, and use the new bucket ranges to determine which rules to display upon expansion.

ask : we should mention this right up front when we're describing the problem.
\end{comment}

\subsection{Using Sum instead of Count}\label{sec:extensions-sum}
Throughout the paper, we define the total score of a rule-list using the marginal counts of rules in the list, and display the count of each rule in our table summary. However, if we have a numerical column (i.e. a `measure' column) in the table, it is straightforward to extend our summary to the `Sum' aggregate over that column instead. Suppose we are given a measure column $c_m$. Then the {\em Sum} for a rule can be defined to be the sum of $c_m$ values over all tuples covered by the rule. {\em MSum} of a rule $r$ in a rule-list $R$ is the sum of $c_m$ values over all tuples covered by $r$ and not covered by any rule in $R$ that occurs before $r$. The Score for $R$ becomes $\text{Score}(R) = \sum_{r\in R} MSum(r,R)W(r)$. Algorithm~\ref{algo:best-rule-set} can be modified to find the best rule set using the new definition of Score, simply by replacing $Count(r)$ by $Sum(r)$ and computing sum and marginal sum instead of count and marginal count in each pass over the table.

### 2014-11-18 TableSummarization/Paper/formal.tex
Addressed Comments. Added NP-Hardness proof

\section{Formal Description} 
We describe our formal problem
in Section~\ref{sec:preliminaries},
describe different scoring functions
in Section~\ref{sec:weighting},
and describe our operator interfaces
in Section~\ref{sec:interface}. 
\subsection{Preliminaries and Definitions}
\label{sec:preliminaries}

@@

Now, onto the second portion: we let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could, as we will see later, 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
on monotonic weighting functions because we prefer 
rules that are more ``specific''
rather than those that are more ``general'' 
(thereby conveying less information).
We further describe our weighting functions in Section~\ref{sec:weighting}.

@@

\stitle{Other Weighting Functions:}
Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
\begin{enumerate}
\item Non-negativity: For all rules $r$, $W(r) \geq 0$.
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\end{enumerate}
A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).

>>>>

\section{Formal Description}\label{sec:formal}
We describe our formal problem
in Section~\ref{sec:preliminaries},
describe different scoring functions
in Section~\ref{sec:weighting},
and describe our operator interfaces
in Section~\ref{sec:interface}. 
\subsection{Preliminaries and Definitions}
\label{sec:preliminaries}

@@

Now, onto the second portion: we let $W$ denote a function that assigns a non-negative {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. 
As we will see, the weighting function does not depend on the specific
tuples in $\calD$, but could, as we will see later, 
depend on the number of $\star$s in $r$,
the schema of $\calD$, 
as well as the number of distinct values in each column of $\calD$.
A weighting function is said to be {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$; we focus
on monotonic weighting functions because we prefer 
rules that are more ``specific''
rather than those that are more ``general'' 
(thereby conveying less information).
We further describe our weighting functions in Section~\ref{sec:weighting}.

@@

\stitle{Other Weighting Functions:}
Even though we have given two example weighting functions here, our algorithms allow the user to set any weighting function $W$, subject to only two conditions:
\squishlist
\item Non-negativity: For all rules $r$, $W(r) \geq 0$.
\item Monotonicity: If $r_1 \geq r_2$, then $W(r_1) \leq W(r_2)$. Monotonicity means that a rule that is less descriptive than another must be assigned a lower weight.
\squishend
A weight function can be used in several ways, including to express higher preference for a column (by assigning higher weight to rules having a non-$\star$ value in that column), or to express indifference to a column (by adding zero weight for having non-$\star$ value in that column).

### 2014-11-18 TableSummarization/Paper/intro.tex
small changes
Addressed Comments. Added NP-Hardness proof

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill down to provide an additional option of anomaly detection.

We further discuss related work in Section~\ref{sec:related}.

@@

\item In Section~\ref{sec:preliminaries}, we introduce the relevant
notation and formally define smart drill down.

\item In Section~\ref{sec:weighting}, we consider different schemes
for weighting rules, along with their practical significance.

\item In Section~\ref{sec:interface}, we describe our
interactive user interface.

>>>>

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}..

Briefly, the contributions of our paper can be summarized are follows: We develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. In addition, we develop a scheme for maintaining and using multiple samples of the table in memory to improve response time on large tables.

@@

\item In Section~\ref{sec:formal}, we introduce the relevant notation and formally define smart drill down. After that, we discuss different schemes for weighting rules, and describe our interactive user interface.
